Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)

### 2022-04-18 doc/manuscript/ESSDmanuscript.Rmd
finish first draft of section 3 (ForC updates)

n_new_variables = 11 + 2 + 2 #delta. + woody.mortality + O horizon
# get n fields modified
sampling_details <- read.csv(paste0(path_to_ForC,"/ForC/database_management_records/record_of_changes.csv"), stringsAsFactors = FALSE, check.names = FALSE)
n_fields_modified =  length(sampling_details[,"Column"]) + 4 
```

@@

Previous versions of ForC [@refs] contained most of the information required by EFDB, and, more broadly, to inform C accounting under IPCC guidelines. 
However, modest changes to the structure and contents of ForC were needed in order to provide all information required by EFDB and to improve ForC's capacity to serve as a repository of valuable information for forest C accounting under IPCC guidelines.
To support export of data to EFDB, and to improve the overall quality of the ForC database, we added or modified `r n_fields_modified` fields (`r table_ForCchanges`), defined `r n_new_variables` new variables, implemented enhanced quality control, manually reviewed ># records to obtain additional required information, and added # new records.

*This section describes changes relative to ForC v2.0 [@anderson-teixeira_forc_2018].*

@@

## Quality control measures
Prior to releasing ForC v4.0, we executed several quality control measures.
First, we implemented a system of continuous integration using GitHub Actions (*sensu* Kim et al. in review) to run some automatic checks any time the master data files are updated.
Second, to improve information on geographic coordinates, we flagged and reviewed records with suspected low precision *(Issue #29)[https://github.com/forc-db/ForC/issues/229]*.
Third, to identify erroneous climate data... *(Issue #212)[https://github.com/forc-db/ForC/issues/212]*.

@@

EFDB data submissions require information that was not recorded in previous versions of ForC.
It was therefore necessary to return to original publications to retrieve relevant information, including the following:

*Data location within source*-  Because EFDB will not accept data which has been digitized (from graphs) while ForC does, it was necessary to review and note the location of the data within the original source

@@

>>>>

n_new_variables = 11 + 2 + 2 #delta. + woody.mortality + O horizon
# get n fields modified
sampling_details <- read.csv(paste0(path_to_ForC,"/ForC/database_management_records/record_of_changes.csv"), stringsAsFactors = FALSE, check.names = FALSE)
n_fields_modified =  length(sampling_details[,"Column"]) + 4 
n_records_checked = NA # this will be hard to get at, but we can at least estimate a minimum (= previously existing records in records sent to EFDB)
n_records_added = NA

```

@@

Previous versions of ForC [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018; @anderson-teixeira_carbon_2021] contained most of the information required by EFDB, and, more broadly, to inform C accounting under IPCC guidelines. 
However, modest changes to the structure and contents of ForC were needed in order to provide all information required by EFDB and to improve ForC's capacity to serve as a repository of valuable information for forest C accounting under IPCC guidelines.
To support export of data to EFDB, and to improve the overall quality of the ForC database, we added or modified `r n_fields_modified` fields (`r table_ForCchanges`), defined `r n_new_variables` new variables, implemented enhanced quality control, manually reviewed ># records to obtain additional required information, and added # new records.

This section describes changes relative to ForC v3.0 [@anderson-teixeira_carbon_2021].

@@

## Quality control measures
Prior to releasing ForC v4.0, we executed several quality control measures.
First, we implemented a system of continuous integration using GitHub Actions [*sensu* @kim_using_inreview] to run some automatic checks any time the master data files are updated, including outlier tests and checks for completeness and naming consistency of records across data files.
Second, to improve information on geographic coordinates, we created a field to record coordinate precision (`r table_ForCchanges`), and flagged and reviewed records with suspected low precision *(Issue #29)[https://github.com/forc-db/ForC/issues/229]*.
Third, to identify erroneous climate data... *(Issue #212)[https://github.com/forc-db/ForC/issues/212]*.

@@

EFDB data submissions required information that was not recorded in previous versions of ForC, but rather was recorded in fields (`r table_ForCchanges`).
It was therefore necessary to return to original publications to retrieve relevant information, including 
(1) whether data were presented in tables / text or digitized from figures (EFDB will not accept digitized data, whereas ForC does), 
(2) estimates in original units, 
and (3) confidence intervals (when not already in ForC).
We also checked that existing ForC records were complete and correct.

Manual review of records was the limiting step for data transfer to EFDB. 
We prioritized review of
(1) records from the Forest Global Earth Observatory [ForestGEO, @anderson-teixeira_ctfsforestgeo_2015; @davies_forestgeo_2021],
(2) studies with confidence intervals recorded in ForC (because uncertainty estimates are important to the IPCC),
(3) original publications containing large numbers of EFDB-relevant records,
and (4) records from tropical regions.
The latter criteria was motivated by the fact that although tropical forest is the single most important biome for climate change mitigation [@refs], ground-based data on tropical forest C cycling tend to be more scarce due to a variety of challenges [@delima_making_2022], and *tropical countries are more likely to apply Tier 1 methodology that bases forest C budgets on previously existing data [@ref].* <!-- I'm almost certain this is true, but need to confirm/ find ref.-->

@@

In addition to reviewing existing records, we added a total of `r n_records_added` new records to ForC.
These included **##** records from **##** studies [@piponiot_distribution_2022; @refs<!--others?-->] that were not previously included in ForC.
In addition, we created new records for **##** EFDB-relevant estimates presented in the original publication that were not yet present in ForC.

### 2022-04-18 doc/manuscript/ESSDmanuscript.tex
finish first draft of section 3 (ForC updates)

Previous versions of ForC \citep{refs} contained most of the information
required by EFDB, and, more broadly, to inform C accounting under IPCC
guidelines. However, modest changes to the structure and contents of
ForC were needed in order to provide all information required by EFDB
and to improve ForC's capacity to serve as a repository of valuable
information for forest C accounting under IPCC guidelines. To support
export of data to EFDB, and to improve the overall quality of the ForC
database, we added or modified 18 fields (Appendix A), defined 15 new
variables, implemented enhanced quality control, manually reviewed
\textgreater\# records to obtain additional required information, and
added \# new records.

\emph{This section describes changes relative to ForC v2.0
\citep{anderson-teixeira_forc_2018}.}

@@

Prior to releasing ForC v4.0, we executed several quality control
measures. First, we implemented a system of continuous integration using
GitHub Actions (\emph{sensu} Kim et al.~in review) to run some automatic
checks any time the master data files are updated. Second, to improve
information on geographic coordinates, we flagged and reviewed records
with suspected low precision \emph{(Issue
\#29){[}https://github.com/forc-db/ForC/issues/229{]}}. Third, to
identify erroneous climate data\ldots{} \emph{(Issue
\#212){[}https://github.com/forc-db/ForC/issues/212{]}}.

@@

EFDB data submissions require information that was not recorded in
previous versions of ForC. It was therefore necessary to return to
original publications to retrieve relevant information, including the
following:

\emph{Data location within source}- Because EFDB will not accept data
which has been digitized (from graphs) while ForC does, it was necessary
to review and note the location of the data within the original source

@@

@@

As of April 16, 2022, ForC (v4.0) contained NA independent records (NA
total), NA of which were for variables relevant to EFDB (Fig. 1). These
records were distributed across all forested continents and ecozones,
with particularly high concentrations in \emph{{[}ecozone(s){]}} and low
concentrations in \emph{{[}ecozone(s){]}} (Fig. 2). The most widely
represented forest type was \emph{{[}type{]}}, followed by
\emph{{[}type{]}} and \emph{{[}type{]}} (Fig. 3).

@@

As of April 16, 2022, we had reviewed and sent NA, NA of which have been
reviewed, accepted, and posted (Fig. 2-3, Table 2). \emph{{[}DETAILS{]}}

>>>>

Previous versions of ForC
\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc_2018, anderson-teixeira_carbon_2021}
contained most of the information required by EFDB, and, more broadly,
to inform C accounting under IPCC guidelines. However, modest changes to
the structure and contents of ForC were needed in order to provide all
information required by EFDB and to improve ForC's capacity to serve as
a repository of valuable information for forest C accounting under IPCC
guidelines. To support export of data to EFDB, and to improve the
overall quality of the ForC database, we added or modified 18 fields
(Appendix A), defined 15 new variables, implemented enhanced quality
control, manually reviewed \textgreater\# records to obtain additional
required information, and added \# new records.

This section describes changes relative to ForC v3.0
\citep{anderson-teixeira_carbon_2021}.

@@

Prior to releasing ForC v4.0, we executed several quality control
measures. First, we implemented a system of continuous integration using
GitHub Actions \citep[\emph{sensu}][]{kim_using_inreview} to run some
automatic checks any time the master data files are updated, including
outlier tests and checks for completeness and naming consistency of
records across data files. Second, to improve information on geographic
coordinates, we created a field to record coordinate precision (Appendix
A), and flagged and reviewed records with suspected low precision
\emph{(Issue \#29){[}https://github.com/forc-db/ForC/issues/229{]}}.
Third, to identify erroneous climate data\ldots{} \emph{(Issue
\#212){[}https://github.com/forc-db/ForC/issues/212{]}}.

@@

EFDB data submissions required information that was not recorded in
previous versions of ForC, but rather was recorded in fields (Appendix
A). It was therefore necessary to return to original publications to
retrieve relevant information, including (1) whether data were presented
in tables / text or digitized from figures (EFDB will not accept
digitized data, whereas ForC does), (2) estimates in original units, and
(3) confidence intervals (when not already in ForC). We also checked
that existing ForC records were complete and correct.

Manual review of records was the limiting step for data transfer to
EFDB. We prioritized review of (1) records from the Forest Global Earth
Observatory
\citep[ForestGEO,][]{anderson-teixeira_ctfsforestgeo_2015, davies_forestgeo_2021},
(2) studies with confidence intervals recorded in ForC (because
uncertainty estimates are important to the IPCC), (3) original
publications containing large numbers of EFDB-relevant records, and (4)
records from tropical regions. The latter criteria was motivated by the
fact that although tropical forest is the single most important biome
for climate change mitigation \citep{refs}, ground-based data on
tropical forest C cycling tend to be more scarce due to a variety of
challenges \citep{delima_making_2022}, and \emph{tropical countries are
more likely to apply Tier 1 methodology that bases forest C budgets on
previously existing data \citep{ref}.}

@@

In addition to reviewing existing records, we added a total of NA new
records to ForC. These included \textbf{\#\#} records from \textbf{\#\#}
studies \citep{piponiot_distribution_2022, refs} that were not
previously included in ForC. In addition, we created new records for
\textbf{\#\#} EFDB-relevant estimates presented in the original
publication that were not yet present in ForC.

@@

As of April 18, 2022, ForC (v4.0) contained NA independent records (NA
total), NA of which were for variables relevant to EFDB (Fig. 1). These
records were distributed across all forested continents and ecozones,
with particularly high concentrations in \emph{{[}ecozone(s){]}} and low
concentrations in \emph{{[}ecozone(s){]}} (Fig. 2). The most widely
represented forest type was \emph{{[}type{]}}, followed by
\emph{{[}type{]}} and \emph{{[}type{]}} (Fig. 3).

@@

As of April 18, 2022, we had reviewed and sent NA, NA of which have been
reviewed, accepted, and posted (Fig. 2-3, Table 2). \emph{{[}DETAILS{]}}
